---
title: "Assignment II: Group [D]"
author: "Ho Chun Hei"
format: 
  html:
    toc: true
    code-fold: false
    embed-resources: true # <= check here: https://quarto.org/docs/reference/formats/html.html
execute:
  echo: true    # <= knitr::opts_chunk$set(echo = TRUE) in Rmarkdown
  warning: false
  message: false
---

# Name

[**Name**]{style="color:red"}**: Ho Chun Hei**

# Generative AI declaration

Students may use generative AI tools during assignment, for example, to check and improve code, provided that the content is original and reflects the student's own contributions. If generative AI is used, students must declare its use in the box below. An example statement is: “*During the assignment, I used \[XXXX\] to \[XXXX\]. I subsequently reviewed and edited the code and take(s) full responsibility for the final version of the code.”*

::: callout
I did not use any generative AI tools for this assignment. All code and analysis are my own original work.
:::

# Dataset

Each group member may choose either of the following dataset:

**Built-in dataset**

-   `palmerpenguins::penguins`
-   `ggplot2::msleep`
-   `ggplot2::midwest`
-   `datasets::quakes`

**Public dataset**

BIOTIME (CC0) <https://biotime.st-andrews.ac.uk/usageGuidelines.php>

-   `BIOTIME_raw_data_192.csv` (uploaded on Canvas) (Whale data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=192>)
-   `BIOTIME_raw_data_232.csv` (uploaded on Canvas) (Fish data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=232>)

# Figures by each group member

### Dataset

::: callout-important
[Which dataset did you choose?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here
ggplot2::midwest
:::

### Introduction

::: callout-important
[What is the dataset about? What variables are included in the data? How was the data taken/measured?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here
This is about counties in the midwest region of the US and the people who live there. For example how many people, their race, no. of school they finished, how many are poor. There are 437 counties in 5 states.

variables: PID, county, state, area, poptotal, popdensity, popwhite, popblack, popamerindian, popasian, popother, percwhite, percblack, percamerindan, percasian, percother, popadults, perchisp, percollege, percprof, percbelowpoverty, percchildbelowpovert, percadultpoverty, percelderlypoverty, inmetro, category

This data came from the United States Census in the 2000. The Census is a big survey that the government does every 10 years. They ask people questions about their age, race, education, and income. This dataset only shows the people in the midwest.
:::

### Interactive plot(s) using `plotly`

::: callout-important
[Please explain which plot type you choose and why. Also, please discuss the strengths and weaknesses of your plot(s).]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here
scatter plot. Because it is a good way to show the relationship 2 variables

One of the strength is it is easy to see patterns with many data points. The hover tooltip is also a good feature that allows us to see the exact values. The legend is clickable that can double click on a category and compare it with the others. We can also single click to hide a category from plot. The zoom feature allows us to drag a box around a specific area and have a closer look. Trend lines can be added to summarize the overall direction of the data.

The weaknesses are when there are many data points, it can become quite crowded and dots may overlap each other. This makes it harder to click on a specific point. Trend lines can show a general pattern can't tell the how strong the correlation is. The plot only shows two variables on the axes at the same time, so other factors are not visible unless use the hover tooltip. Finally, it must be viewed on a computer for the full experience.

:::

Note: At least one plot is required, but multiple plots are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

To see if counties with more college graduates have less poverty. I want to know if going to college helps people make more money or not.
:::

Note: If you want to specify setting for a code chunk, you can use `#|` for each option (for example, `#| echo: true`). Rmarkdown code chunk still works (for example, `{r echo = true}`), but Quarto style is recommended.

```{r}
#| echo: true
# YOUR CODE HERE
# a scatter plot showing education vs poverty with hover tooltips and trend lines to show their relationships
library(tidyverse)
library(plotly)

p <- ggplot(midwest, aes(x = percollege, y = percbelowpoverty, color = state)) +
  geom_point(alpha = 0.7, size = 2) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) + # Add trend line 
  labs(
    title = "College Education vs Poverty in the Midwest", # Add title and axis
    x = "percentage with college degree",
    y = "percentage below entry"
  ) +
  theme_minimal()

ggplotly(p)
```
```
```

### Animation using `gganimate`

::: callout-important
[Please explain which animation type you choose and why. Also, please discuss the strengths and weaknesses of your animation.]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here
My animation switches between the five states. I chose this because my data does not have a time variable so instead of showing the change over time, I show how the pattern looks in different states.

One of the strengths is it shows only one state at one time so its not confusing. The dots fade in and out making the animation looks a bit smoother. The state name also changes in the title with the animation

The weaknesses of this animation are that it loops over and over again non stop. To see a specific state, we have to wait for it to come back. It also does not work on printed paper because it is animated.
:::

Note: At least one animation is required, but multiple animations are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

To show how the relationship between college education and poverty in the five states. Watching the states cycle can compare the patterns and see that some states have a stronger connection between education and poverty than others.
:::

```{r}
# animated scatter plot that repeats the 5 states to compare the education and poverty patterns
library(ggplot2)
library(gganimate)
library(gifski)
data("midwest")

p_anim <- ggplot(midwest, aes(x = percollege, y = percbelowpoverty, 
                              size = poptotal/1000, color = state)) +
  geom_point(alpha = 0.7) +
  scale_color_brewer(palette = "Set1", guide = "none") +
  scale_size_continuous(name = "Population (thousands)", range = c(2, 10)) +
  labs(
    title = "State: {closest_state}", # Add title and axis
    x = "percentage with college degree",
    y = "percentage below entry"
  ) +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(face = "bold", size = 18))

anim <- p_anim + # Add animation to the plot
  transition_states(state, transition_length = 3, state_length = 2) +
  ease_aes('cubic-in-out') +
  enter_fade() + # Makes dots fade in when appearing
  exit_fade() + # Makes dots shrink when disappearing
  shadow_wake(wake_length = 0.2)

animate(anim, nframes = 100, fps = 10, width = 700, height = 500) # Make the animation as a GIF
```

