---
title: "Assignment II: Group D"
author: "Chau Yat Nam"
format: 
  html:
    toc: true
    code-fold: false
    embed-resources: true 
execute:
  echo: true
  warning: false
  message: false
---

# Name

[**Name**]{style="color:red"}**: Chau Yat Nam**

# Generative AI declaration

Students may use generative AI tools during assignment, for example, to check and improve code, provided that the content is original and reflects the student's own contributions. If generative AI is used, students must declare its use in the box below. An example statement is: “*During the assignment, I used \[XXXX\] to \[XXXX\]. I subsequently reviewed and edited the code and take(s) full responsibility for the final version of the code.”*

::: callout
During the assignment, I used Grok AI to help modify code and provide structure suggestions for the Quarto document. I reviewed, tested, and edited all code and text myself and take full responsibility for the final version.
:::

# Dataset

Each group member may choose either of the following dataset:

**Built-in dataset**

-   `palmerpenguins::penguins`
-   `ggplot2::msleep`
-   `ggplot2::midwest`
-   `datasets::quakes`

**Public dataset**

BIOTIME (CC0) <https://biotime.st-andrews.ac.uk/usageGuidelines.php>

-   `BIOTIME_raw_data_192.csv` (uploaded on Canvas) (Whale data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=192>)
-   `BIOTIME_raw_data_232.csv` (uploaded on Canvas) (Fish data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=232>)

# Figures by each group member

### Dataset

::: callout-important
[Which dataset did you choose?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

I chose the built-in dataset `palmerpenguins::penguins`.
:::

### Introduction

::: callout-important
[What is the dataset about? What variables are included in the data? How was the data taken/measured?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

The `palmerpenguins` dataset contains size measurements for **344 adult penguins** of three species (Adélie, Chinstrap, and Gentoo) observed on three islands (Biscoe, Dream, Torgersen) in the Palmer Archipelago, Antarctica.  

**Key variables** include:  
- `species`, `island`, `sex`, `year` (categorical)  
- `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g` (numeric measurements)  

The data were collected by Dr. Kristen Gorman and the Palmer Station Long Term Ecological Research (LTER) program between 2007–2009.
:::

### Interactive plot(s) using `plotly`

::: callout-important
[Please explain which plot type you choose and why. Also, please discuss the strengths and weaknesses of your plot(s).]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

I chose an interactive scatterplot (created with `ggplot2` + `ggplotly`).  
**Why?** It clearly shows the strong positive relationship between flipper length and body mass while allowing users to hover for exact values, filter by species, and zoom in.  

**Strengths**: High interactivity (hover, zoom, legend toggle), easy to explore multivariate data, publication-ready when rendered to HTML.  
**Weaknesses**: Can become cluttered with very large datasets; hover labels must be carefully designed to avoid information overload.
:::

Note: At least one plot is required, but multiple plots are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

The purpose is to communicate that **Gentoo penguins are distinctly larger** than the other two species and that flipper length and body mass are strongly correlated across all species. This helps the reader quickly understand species-level differences in body size.
:::

Note: If you want to specify setting for a code chunk, you can use `#|` for each option (for example, `#| echo: true`). Rmarkdown code chunk still works (for example, `{r echo = true}`), but Quarto style is recommended.

```{r}
#| echo: true

# Install the package automatically if it's not already installed
if (!requireNamespace("palmerpenguins", quietly = TRUE)) {
  install.packages("palmerpenguins")
}

library(palmerpenguins)
library(ggplot2)
library(plotly)
library(dplyr)
library(tidyr)

penguins_clean <- penguins |> 
  drop_na()

p <- ggplot(penguins_clean, aes(x = flipper_length_mm, 
                                y = body_mass_g, 
                                color = species,
                                text = paste("Species:", species,
                                             "<br>Island:", island,
                                             "<br>Sex:", sex))) +
  geom_point(size = 2, alpha = 0.8) +
  labs(title = "Flipper Length vs Body Mass by Penguin Species",
       x = "Flipper Length (mm)",
       y = "Body Mass (g)") +
  theme_minimal()

ggplotly(p, tooltip = "text")
```

### Animation using `gganimate`

::: callout-important
[Please explain which animation type you choose and why. Also, please discuss the strengths and weaknesses of your animation.]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

I chose a `transition_time()` animation that reveals changes across the three observation years.  
**Why?** It adds a temporal dimension and makes the static relationship more engaging.  

**Strengths**: Dynamic and memorable; viewers can see how measurements evolve over time.  
**Weaknesses**: The time span is only 3 years, so the animation is short; requires the `gganimate` and `gifski` packages to render properly.
:::

Note: At least one animation is required, but multiple animations are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

The purpose is to show that **the positive relationship between body mass and flipper length holds consistently across years**, while highlighting that Gentoo penguins remain the largest group throughout the study period.
:::

```{r}
library(gganimate)

anim <- ggplot(penguins_clean, aes(x = flipper_length_mm, 
                                   y = body_mass_g, 
                                   color = species)) +
  geom_point(size = 2, alpha = 0.7) +
  labs(title = "Penguin Size Over Time: Year {frame_time}",
       x = "Flipper Length (mm)",
       y = "Body Mass (g)") +
  theme_minimal() +
  transition_time(year) +
  ease_aes("linear")

animate(anim, nframes = 100, fps = 10, width = 800, height = 500)
```
