---
title: "Assignment II: GroupD "
author: "Chan Tsz Ting"
format: 
  html:
    toc: true
    code-fold: false
    embed-resources: true # <= check here: https://quarto.org/docs/reference/formats/html.html
execute:
  echo: true    # <= knitr::opts_chunk$set(echo = TRUE) in Rmarkdown
  warning: false
  message: false
---

# Name

[**Name**]{style="color:red"}**: Chan Tsz Ting**

# Generative AI declaration

Students may use generative AI tools during assignment, for example, to check and improve code, provided that the content is original and reflects the student's own contributions. If generative AI is used, students must declare its use in the box below. An example statement is: “*During the assignment, I used \[XXXX\] to \[XXXX\]. I subsequently reviewed and edited the code and take(s) full responsibility for the final version of the code.”*

::: callout
xxxxxxxxxxxxxx
:::

# Dataset

Each group member may choose either of the following dataset:

**Built-in dataset**

-   `palmerpenguins::penguins`
-   `ggplot2::msleep`
-   `ggplot2::midwest`
-   `datasets::quakes`

**Public dataset**

BIOTIME (CC0) <https://biotime.st-andrews.ac.uk/usageGuidelines.php>

-   `BIOTIME_raw_data_192.csv` (uploaded on Canvas) (Whale data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=192>)
-   `BIOTIME_raw_data_232.csv` (uploaded on Canvas) (Fish data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=232>)

# Figures by each group member

### Dataset

::: callout-important
[Which dataset did you choose?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here
datasets::quakes

xxxxxxxxxxxxxx
:::

### Introduction

::: callout-important
[What is the dataset about? What variables are included in the data? How was the data taken/measured?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here
The dataset contains 1,000 seismic events (earthquakes) occurring near Fiji since 1964. It includes the following variables.lat or long which is the geographic coordinates of the event.
Depth which is the depth of the earthquake in kilometers (km).Mag which is the Richter magnitude of the earthquake.And finally the stations which is the number of stations that reported the event.

xxxxxxxxxxxxxx
:::

### Interactive plot(s) using `plotly`

::: callout-important
[Please explain which plot type you choose and why. Also, please discuss the strengths and weaknesses of your plot(s).]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here
I choose a 3D Scatter Plot using lat, long, and depth because earthquakes happen in 3D space. A static plot makes it hard to see the relationship between location and depth, but an interactive 3D plot allows the user to rotate the view to see the subduction zone.The strengths of it is allows for precise exploration of data points through hovering and rotation. While the weaknesses is 3D plots can sometimes be difficult to interpret on a 2D screen without active user interaction.

xxxxxxxxxxxxxx
:::

Note: At least one plot is required, but multiple plots are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here
There are few purposes. First, Accurate Spatial Representation.Physical Context like Earthquakes occur at various depths, and a 3D plot allows us to map depth to the Z-axis, representing the actual underground location.and Structural Clarity helps the reader visualize the "story" of the data, such as the angle of a subduction zone where one tectonic plate slides beneath another.

Second, Multi-Variable Integration.We can display four variables simultaneously—longitude, latitude, depth, and magnitude—without making the chart look cluttered.And by using color or size for magnitude, the visualization has a focused "target," allowing the reader to immediately see if larger earthquakes happen at specific depths.

xxxxxxxxxxxxxx
:::

Note: If you want to specify setting for a code chunk, you can use `#|` for each option (for example, `#| echo: true`). Rmarkdown code chunk still works (for example, `{r echo = true}`), but Quarto style is recommended.

```{r}
#| echo: true
# YOUR CODE HERE
library(plotly)

# Create a 3D scatter plot of earthquake locations
plot_ly(quakes, 
        x = ~long, y = ~lat, z = ~-depth, # Negative depth to show it underground
        color = ~mag, 
        size = ~mag,
        text = ~paste("Magnitude:", mag, "<br>Depth:", depth, "km"),
        type = "scatter3d", 
        mode = "markers") %>%
  layout(title = "3D Mapping of Fiji Earthquakes",
         scene = list(xaxis = list(title = 'Longitude'),
                      yaxis = list(title = 'Latitude'),
                      zaxis = list(title = 'Depth (km)')))
```

### Animation using `gganimate`

::: callout-important
[Please explain which animation type you choose and why. Also, please discuss the strengths and weaknesses of your animation.]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here
I choose A transitioning scatter plot (using transition_states or transition_filter) as by using transition_states to cycle through different depth levels,i can show how the relationship between magnitude and reporting stations changes as i go deeper into the Earth. Also, the quakes dataset has 1,000 points. Animating them in "chunks" (by depth or magnitude) prevents the screen from becoming a cluttered mess of overlapping dots, making it easier for the reader to follow.For strengths, it emphasizes the "story" that stronger earthquakes are detected by significantly more stations.But for weaknesses, animations can be distracting if the transitions are too fast or if there is too much data moving at once.

xxxxxxxxxxxxxx
:::

Note: At least one animation is required, but multiple animations are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here
The purpose of this animation is to demonstrate that while larger magnitude earthquakes are generally detected by more stations, this relationship is also influenced by the depth of the seismic event.

xxxxxxxxxxxxxx
:::

```{r}
# YOUR CODE HERE
library(ggplot2)
library(gganimate)

# Plotting Magnitude vs Stations, animated by Depth categories
ggplot(quakes, aes(x = mag, y = stations, color = mag)) +
  geom_point(alpha = 0.7, size = 3) +
  theme_minimal() +
  labs(title = 'Magnitude vs Reporting Stations',
       subtitle = 'Showing Depth: {closest_state} km',
       x = 'Richter Magnitude',
       y = 'Number of Stations') +
  transition_states(cut(depth, breaks = 4), 
                    transition_length = 2, 
                    state_length = 1) +
  ease_aes('linear')
```
